---
title: "DSS Computing Bootcamp Template"
author: "Mine Cetinkaya-Rundel"
date: "8/21/2015"
output: html_document
---

## Load packages

Load the following packages for 

- data wrangling (`dplyr`), and 
- data visualizaiton (`ggplot2`)

```{r message=FALSE}
library(dplyr)
library(ggplot2)
```

## Load data

```{r}
gap <- read.csv("https://bit.ly/gap_data_upd")
```

## Exercises

### Exercise 1

How many observations are in this dataset?

```{r nrow, echo = FALSE}
nrow(gap)
```

There are `r nrow(gap)` observations in the database.

### Exercise 2

Visualize the relationship between GDP and life expectancy for 
countries in Europe in 1952 using a scatter plot.

```{r eu_52}
eu_52 <- gap %>% 
  filter(continent == "Europe", year == 1952)
ggplot(data = eu_52, aes(x = gdpPercap, y = lifeExp)) +
  geom_point()  
```

### Exercise 3

Add year 1967 in another color.

```{r eu_52_67}
eu_52 <- gap %>% 
  filter(continent == "Europe", year == 1952)
eu_67 <- gap %>% 
  filter(continent == "Europe", year == 1967)
ggplot(data = eu_52, aes(x = gdpPercap, y = lifeExp),color = factor(1952)) +
  geom_point() + geom_point(data = eu_67, color = factor(1967))

eu_5267 <- gap %>% 
filter(continent == "Europe", year %in% c(1952,1967))# year==1952|year==1967
ggplot(data = eu_5267, aes(x = gdpPercap, y = lifeExp,color = factor(year))) + geom_point()#color='blue'          # remove facor and make it as numerical 

```

